---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "")
library(RsimdDispatch)
```
# simdDispatch

A small, self-contained C library for runtime SIMD dispatch. The build
system probes the compiler for each ISA at configure time, compiles matching
kernel objects with the appropriate flags, and links them into a static archive.
At runtime, CPU feature detection selects the best available backend; the caller
invokes an operation through a single `sd_dispatch_invoke()` call and the library
routes it to the right kernel.

`simd_dispatch.c`, `cpu_features.c`, and the scalar kernel have no external
dependencies beyond a C11 compiler. SIMDe is a dependency of the SIMD backend
kernels only (`kernels/kernel_sse2.c`, `kernel_avx2.c`, etc.) — the dispatch
core and the scalar path are completely independent of it. If SIMDe headers are
absent the build falls back to a scalar-only library.

## Building

```bash
cd tools/simdDispatch
make          # builds build/libsimd_dispatch.a
make test     # builds and runs the test suite
make examples # builds example binaries
make info     # shows what the compiler supports on this machine
```

`SIMDE_DIR` can be overridden if SIMDe headers are not at `../../inst/include`:

```bash
make SIMDE_DIR=/path/to/simde-parent test
```

## What the compiler sees on this machine

```{r info, engine="bash", engine.opts=list(bash="-l")}
cd "$(git rev-parse --show-toplevel)/tools/simdDispatch" && make info 2>/dev/null
```

## Test results

```{r tests, engine="bash", engine.opts=list(bash="-l")}
cd "$(git rev-parse --show-toplevel)/tools/simdDispatch" && make test 2>/dev/null
```

## API

The library is initialised once and then used through three calls:

```c
sd_init_dispatch();   /* auto-select the best compiled + available backend */
sd_set_backend("avx2");   /* or pick explicitly; "auto" reverts */

SdCountNonzeroCall call = { .x = buf, .n = n, .result = 0 };
sd_dispatch_invoke(SD_OP_COUNT_NONZERO, SD_SIG_RAW_COUNT, &call, "count_nonzero");
```

`sd_dispatch_invoke` writes the result into the call-frame struct; there is no
return value and no allocation. `sd_init_dispatch` is idempotent and
thread-safe for concurrent reads after the first call.

The error handler is pluggable. The default writes to stderr and calls
`abort()`; the R adapter installs one that redirects to `Rf_error()`.

## Operations

The two operations compiled into the reference kernels are
`SD_OP_COUNT_NONZERO` (`uint8_t` buffer → `size_t`) and `SD_OP_CONVOLVE1D`
(`double` arrays → `double` output). Adding an operation requires entries in
`kernels/kernel_api.h` (operation enum value, signature enum value, call-frame
struct), a kernel function per backend, and a `SdKernelDef` table registered
with `sd_register_kernel_table()`.

## Runtime diagnostics (via the R package)

```{r diag}
info <- simd_info()
info[c("compiled_backends", "simde_native_backends",
       "available_backends", "target_arch", "target_os")]
```

## Benchmarks

Timings on this machine comparing scalar against the best available backend
for the two built-in operations.

```{r bench, message=FALSE}
if (requireNamespace("bench", quietly = TRUE)) {
  x <- rep(as.raw(c(0L, 1L, 2L, 3L, 0L, 255L, 7L, 0L)), length.out = 2^20)
  best <- tail(simd_info()$available_backends, 1)

  b1 <- bench::mark(
    scalar = { simd_set_backend("scalar"); count_nonzero(x) },
    best   = { simd_set_backend(best);   count_nonzero(x) },
    iterations = 20, check = TRUE
  )
  simd_set_backend("auto")
  print(b1[, c("expression", "median", "itr/sec")])

  a <- runif(2^14); kernel <- runif(64)
  b2 <- bench::mark(
    scalar = { simd_set_backend("scalar"); convolve1d(a, kernel) },
    best   = { simd_set_backend(best);   convolve1d(a, kernel) },
    iterations = 20, check = TRUE, relative = FALSE
  )
  simd_set_backend("auto")
  print(b2[, c("expression", "median", "itr/sec")])
} else {
  message("install the 'bench' package to see timings")
}
```
