---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# RsimdDispatch

<!-- badges: start -->
[![R-CMD-check](https://github.com/sounkou-bioinfo/RsimdDispatch/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sounkou-bioinfo/RsimdDispatch/actions/workflows/R-CMD-check.yaml)
[![R-universe](https://sounkou-bioinfo.r-universe.dev/badges/RsimdDispatch)](https://sounkou-bioinfo.r-universe.dev/RsimdDispatch)
<!-- badges: end -->

Pure-C runtime SIMD dispatch templates for R packages. Compile several kernel
variants, then select only a backend that is both compiled and supported by the
current CPU.

```r
simd_set_backend("scalar")
count_nonzero(x)

simd_set_backend("avx2")
count_nonzero(x)

simd_set_backend("auto")
```

## Install

```r
install.packages(
  "RsimdDispatch",
  repos = c("https://sounkou-bioinfo.r-universe.dev", "https://cloud.r-project.org")
)
```

From GitHub:

```r
# install.packages("remotes")
remotes::install_github("sounkou-bioinfo/RsimdDispatch")
```

From a checkout:

```sh
R CMD INSTALL .
```

## API

```r
count_nonzero(x)          # demo C kernel over raw vectors
simd_set_backend("auto")  # auto, scalar, sse2, sse41, avx2, avx512, neon
simd_backend()            # selected backend
simd_info()               # compiled/supported/selected backend info
use_simd_dispatch(path)   # copy the C template into another package
```

`simd_set_backend()` rejects uncompiled or CPU-unsupported backends. There is no
unsafe override.

## Demo

```{r}
library(RsimdDispatch)

x <- as.raw(c(0, 1, 2, 0, 255))
count_nonzero(x)

simd_backend()
simd_info()[c("compiled_backends", "cpu_supported_backends", "available_backends")]
```

Switching backends is same-process because all variants are linked into one R
shared library and selected through guarded function pointers.

```{r}
simd_set_backend("scalar")
scalar_count <- count_nonzero(x)

simd_set_backend("avx2")
avx2_count <- count_nonzero(x)

simd_set_backend("auto")
data.frame(
  scalar = scalar_count,
  avx2 = avx2_count,
  selected_after_auto = simd_backend()
)
```

## Scalar vs AVX2 benchmark

```{r, echo = FALSE, results = "asis"}
set.seed(1)
n <- 50 * 1024 * 1024
x <- as.raw(sample.int(256L, n, replace = TRUE) - 1L)
reps <- 20L

bench_backend <- function(backend, x, reps) {
  simd_set_backend(backend)
  invisible(count_nonzero(x))
  invisible(gc())
  elapsed <- system.time({
    ans <- 0
    for (i in seq_len(reps)) {
      ans <- ans + count_nonzero(x)
    }
  })[["elapsed"]]
  data.frame(
    backend = simd_backend(),
    scanned_mb = length(x) * reps / 1e6,
    seconds = elapsed,
    mb_per_second = length(x) * reps / 1e6 / elapsed,
    checksum = ans,
    row.names = NULL
  )
}

bench <- rbind(
  bench_backend("scalar", x, reps),
  bench_backend("avx2", x, reps)
)
stopifnot(length(unique(bench$checksum)) == 1L)
bench$speedup_vs_scalar <- bench$seconds[bench$backend == "scalar"] / bench$seconds
simd_set_backend("auto")
knitr::kable(bench[c("backend", "scanned_mb", "seconds", "mb_per_second", "speedup_vs_scalar")], digits = 3)
```

## Vendoring flow

SIMDe is vendored as full header tree under `inst/include/simde`. Downstream
packages can use:

```text
LinkingTo: RsimdDispatch
```

and include headers such as:

```c
#include <simde/x86/avx2.h>
#include <simde/arm/neon.h>
```

Update the vendored checkout and generated notices with:

```sh
Rscript tools/vendor-simde.R
Rscript tools/update-authors.R
```

Provenance lives in `inst/vendor/simde/VERSION`, `inst/AUTHORS`, and
`inst/LICENCE.note`.

## Development

The top-level Makefile uses GNU make.

```sh
make rd       # roxygen docs and NAMESPACE
make readme   # evaluate README.Rmd and write README.md
make test
make check
```
