---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# RsimdDispatch

<!-- badges: start -->
[![R-CMD-check](https://github.com/sounkou-bioinfo/RsimdDispatch/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sounkou-bioinfo/RsimdDispatch/actions/workflows/R-CMD-check.yaml)
[![R-universe](https://sounkou-bioinfo.r-universe.dev/badges/RsimdDispatch)](https://sounkou-bioinfo.r-universe.dev)
<!-- badges: end -->

Pure-C runtime SIMD dispatch templates for R packages. Stage scalar, SSE,
AVX, AVX-512, NEON, and WebAssembly SIMD128 kernel objects during
configuration; link one shared library and switch among compiled and
CPU-supported implementations at runtime.

## Install

```r
install.packages(
  "RsimdDispatch",
  repos = c("https://sounkou-bioinfo.r-universe.dev", "https://cloud.r-project.org")
)
```

From GitHub:

```r
# install.packages("remotes")
remotes::install_github("sounkou-bioinfo/RsimdDispatch")
```

## Use in other packages

Copy the C dispatch scaffold from your package root:

```r
RsimdDispatch::use_simd_dispatch(pkg = "MyPackage", prefix = "mypkg")
```

The helper updates `DESCRIPTION` with `LinkingTo: RsimdDispatch`, writes the
scaffold files, and substitutes package-specific registration and C symbol
prefixes. It does not add a runtime dependency on `RsimdDispatch`. Replace the demo `count_nonzero()` and `convolve1d()` kernels under `tools/kernels/` with your own
scalar/SSE/AVX/NEON/WebAssembly kernels. The generated `configure` script probes
compiler support, stages selected kernel objects in `src/rsd-kernels/`, and leaves
`src/Makevars` to link those objects with the baseline R API, CPU detection, and
dispatcher.

See the package vignettes for the downstream-package workflow and dispatch
semantics.

## API

```{r}
library(RsimdDispatch)

x <- as.raw(c(0, 1, 2, 0, 255))
a <- c(1, 2, 3)
b <- c(10, 100)

count_nonzero(x)
convolve1d(a, b)
simd_backend()
simd_set_backend("scalar")
simd_backend()
simd_set_backend("avx2")
simd_backend()
simd_set_backend("auto")
simd_info()[c("compiled_backends", "cpu_supported_backends", "available_backends", "simde_native_backends")]
simde_info()[c("version", "commit")]
simd_dispatch_template_path()
```

`simd_set_backend()` rejects uncompiled or CPU-unsupported backends.

## Example

```{r}
count_nonzero(x)
convolve1d(a, b)
simd_info()[c("selected_backend", "available_backends")]
```

Switching backends is same-process because all variants are linked into one R
shared library and selected through a guarded operation table.

```{r}
simd_set_backend("scalar")
scalar_count <- count_nonzero(x)
scalar_conv <- convolve1d(a, b)

simd_set_backend("avx2")
avx2_count <- count_nonzero(x)
avx2_conv <- convolve1d(a, b)

simd_set_backend("auto")
data.frame(
  scalar = scalar_count,
  avx2 = avx2_count,
  convolution_equal = identical(scalar_conv, avx2_conv),
  selected_after_auto = simd_backend()
)
```

## Scalar vs AVX2 benchmarks

```{r, results = "asis"}
set.seed(1)
x <- as.raw(sample.int(256L, 50 * 1024 * 1024, replace = TRUE) - 1L)

mark <- bench::mark(
  scalar = {
    simd_set_backend("scalar")
    count_nonzero(x)
  },
  avx2 = {
    simd_set_backend("avx2")
    count_nonzero(x)
  },
  iterations = 20,
  check = TRUE,
  memory = FALSE
)

simd_set_backend("auto")

bench <- data.frame(
  backend = as.character(mark$expression),
  median_ms = as.numeric(mark$median) * 1000,
  mb_per_second = length(x) * as.numeric(mark$`itr/sec`) / 1e6,
  iterations = mark$n_itr,
  row.names = NULL
)
bench$speedup_vs_scalar <- bench$mb_per_second / bench$mb_per_second[bench$backend == "scalar"]

knitr::kable(bench, digits = 3)
```

The same runtime switch can benchmark a full one-dimensional convolution: the
classic nested-loop `out[i + j - 1] += a[i] * b[j]` kernel.

```{r, results = "asis"}
set.seed(2)
a <- runif(100000)
b <- runif(100)

conv_mark <- bench::mark(
  scalar = {
    simd_set_backend("scalar")
    convolve1d(a, b)
  },
  avx2 = {
    simd_set_backend("avx2")
    convolve1d(a, b)
  },
  iterations = 20,
  check = TRUE,
  memory = FALSE
)

simd_set_backend("auto")

conv_bench <- data.frame(
  backend = as.character(conv_mark$expression),
  median_ms = as.numeric(conv_mark$median) * 1000,
  million_multiply_adds_per_second = (length(a) * length(b)) * as.numeric(conv_mark$`itr/sec`) / 1e6,
  iterations = conv_mark$n_itr,
  row.names = NULL
)
conv_bench$speedup_vs_scalar <- conv_bench$million_multiply_adds_per_second /
  conv_bench$million_multiply_adds_per_second[conv_bench$backend == "scalar"]

knitr::kable(conv_bench, digits = 3)
```

## Development

### Updating bundled SIMDe

This is a maintainer workflow for `RsimdDispatch` itself, not something
downstream packages need to run. Downstream packages use `LinkingTo:
RsimdDispatch` and the copied dispatch template; they do not re-vendor SIMDe.

`RsimdDispatch` vendors the full SIMDe header tree under `inst/include/simde`.
To update the pinned checkout and regenerate bundled-code notices:

```sh
Rscript tools/vendor-simde.R
Rscript tools/update-authors.R
```

Provenance lives in `inst/vendor/simde/VERSION`, `inst/AUTHORS`, and
`inst/LICENCE.note`.

### Package maintenance

The top-level Makefile uses GNU make.

```sh
make rd       # roxygen docs and NAMESPACE
make readme   # evaluate README.Rmd and write README.md
make test
make check
bash tools/webr-build-check.sh  # build and check the webR/WebAssembly package
```
