---
title: "Getting Started with RsimdDispatch"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with RsimdDispatch}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`RsimdDispatch` demonstrates a single-shared-library runtime dispatch pattern
for C code in R packages. The package stages scalar and SIMD kernel objects
during configuration and switches among compiled, CPU-supported implementations
through a guarded resolved operation table.

```{r}
library(RsimdDispatch)

x <- as.raw(c(0, 1, 2, 0, 255))
a <- c(1, 2, 3)
b <- c(10, 100)
count_nonzero(x)
convolve1d(a, b)
simd_backend()
```

Inspect the installed build:

```{r}
simd_info()[c("compiled_backends", "available_backends", "operation_selected_backends")]
```

Switching is allowed in the same R process:

```{r}
simd_set_backend("scalar")
count_nonzero(x)
convolve1d(a, b)

candidate <- setdiff(simd_info()$operation_backends$convolve1d, "scalar")[1]
if (!is.na(candidate)) {
  simd_set_backend(candidate)
  count_nonzero(x)
  convolve1d(a, b)
}

simd_set_backend("auto")
simd_backend()
```

An explicit backend is accepted only when it was compiled and the current
CPU/runtime supports it. Unsupported choices error before any ISA-specific code
runs.
