---
title: "Runtime Dispatch Semantics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Runtime Dispatch Semantics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`RsimdDispatch` uses a single R shared library. All compiled variants are linked
into that shared object, and a function pointer selects the active
implementation.

This means backend switching is safe in one R process:

```{r}
library(RsimdDispatch)
x <- as.raw(c(0, 1, 2, 3))

simd_set_backend("scalar")
count_nonzero(x)

candidate <- setdiff(simd_info()$available_backends, "scalar")[1]
if (!is.na(candidate)) {
  simd_set_backend(candidate)
  count_nonzero(x)
}

simd_set_backend("auto")
simd_backend()
```

`simd_set_backend()` performs two checks for explicit choices:

- the backend was compiled into this package build;
- the current CPU/runtime supports the backend, including OS support for AVX and
  AVX-512 register state on x86.

There is intentionally no unsafe force mode. If a backend is not available, the
setter errors before any SIMD-only instruction can execute.

```{r}
simd_info()[c("compiled_backends", "cpu_supported_backends", "available_backends")]
```

`"auto"` selects the best backend from the compiled and supported intersection.
The current ranking is:

```text
avx512 > avx2 > sse41 > sse2 > neon > scalar
```

Architecture guards mean x86 systems normally consider x86 backends and ARM
systems normally consider NEON. Scalar is always available.
