---
title: "Using RsimdDispatch in Other Packages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using RsimdDispatch in Other Packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Downstream packages use `RsimdDispatch` as a template and header provider. The
usual workflow is:

1. call `use_simd_dispatch()` from the package root;
2. replace the demo `count_nonzero()` and `convolve1d()` kernels with package-specific kernels;
3. keep CPU detection, dispatch, and R API files compiled by R's ordinary `src/Makevars` path;
4. let `configure` stage scalar and ISA-specific kernel objects before linking one shared library.

`use_simd_dispatch()` updates `DESCRIPTION`, adding `RsimdDispatch` to
`LinkingTo`, and copies package-specific scaffold files. It does not add a
runtime dependency on `RsimdDispatch`. The `LinkingTo` entry makes bundled SIMDe
headers available to C code:

```c
#include <simde/x86/avx2.h>
#include <simde/arm/neon.h>
```

## Copy the scaffold

From the downstream package root:

```{r, eval = FALSE}
RsimdDispatch::use_simd_dispatch(pkg = "MyPackage", prefix = "mypkg")
```

If `pkg` is omitted, the package name is read from `DESCRIPTION`. If `prefix` is
omitted, a lowercase C identifier prefix is derived from the package name. The
helper performs the important substitutions, including `@useDynLib`,
`R_init_<Package>()`, `rsd_`, `RSD_`, and `RC_` symbols.

The copied scaffold includes:

```text
configure
configure.win
cleanup
tools/configure-simd-dispatch.sh
tools/kernels/kernel_scalar.c
tools/kernels/kernel_sse2.c
tools/kernels/kernel_sse41.c
tools/kernels/kernel_avx2.c
tools/kernels/kernel_avx512.c
tools/kernels/kernel_neon.c
tools/kernels/kernel_wasm_simd128.c
tools/kernels/kernel_common.h
src/Makevars.in
src/Makevars.win.in
src/cpu_features.c
src/simd_dispatch.c
```

## Replace the demo kernels

The demo operations are registered by each backend file. For example, the AVX2
staged source registers both demo operations:

```c
void rsd_register_avx2(RsdDispatchBuilder *builder) {
    rsd_register_count_nonzero(builder, rsd_count_nonzero_avx2);
    rsd_register_convolve1d(builder, rsd_convolve1d_avx2);
}
```

SSE2 and SSE4.1 currently register only `count_nonzero()`, which demonstrates
operation-level backend support. A backend intentionally omits an operation by
not registering that slot. Explicitly selecting such a backend is allowed, but
calling the unsupported operation errors clearly; `"auto"` skips that backend
for the unsupported operation.

For a real package, replace the demo operation signatures with your own and keep
the same structure:

```text
R API wrapper        ordinary src/Makevars compilation
CPU feature checks   ordinary src/Makevars compilation
dispatch table       ordinary src/Makevars compilation
scalar kernel        staged by configure under src/rsd-kernels/
SSE/AVX/NEON/wasm files   staged by configure as optional objects under src/rsd-kernels/
```

Do not put `-mavx2`, `-mavx512*`, `-msimd128`, or `-march=native` in global package flags.
The configure helper keeps ISA flags local to optional staged kernel objects, and
the dispatcher remains safe on baseline CPUs.

## Extending the number of operations

The staged build model already scales across operations: if you add another
function to each existing `tools/kernels/kernel_*.c` file, the same staged object
for that backend contains the new implementation. For a small number of
operations, keep the dispatch code explicit and add:

1. a function-pointer typedef and typed registration helper in
   `src/simd_dispatch.h` and `src/simd_dispatch.c`;
2. one implementation per backend file under `tools/kernels/`, following the
   naming convention `rsd_<operation>_<backend>()`;
3. a call to the typed registration helper in each backend file that implements
   the operation;
4. the resolved dispatch slot and wrapper in `src/simd_dispatch.c`;
5. the `.Call` wrapper in `src/r_api.c`;
6. the native registration entry in `src/registration.c`;
7. the R wrapper, tests, and documentation.

Backend metadata is table-driven, while operation implementation ownership stays
close to each backend source file. This explicit layout is easier to read for
the template and still keeps backend-selection logic centralized.

For packages with many public operations, keep a single operation catalog that
can generate the repetitive dispatch metadata, `.Call` registration, and R
wrappers. Optimized backend kernels can remain hand-written, but operation names,
argument lists, and exported wrapper metadata should not be duplicated by hand.

## Build

The generated `configure` checks compiler support, writes `src/config.h` plus
`src/Makevars`, and stages selected kernel objects in `src/rsd-kernels/`. Runtime CPU
support is checked later by the installed package:

```sh
R CMD INSTALL .
```
