---
title: "Using RsimdDispatch in Other Packages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using RsimdDispatch in Other Packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Downstream packages use `RsimdDispatch` as a template and header provider. The
usual workflow is:

1. add `RsimdDispatch` to `LinkingTo`;
2. copy the dispatch scaffold;
3. replace the demo `count_nonzero()` kernels with package-specific kernels;
4. keep CPU detection and dispatch files compiled with baseline flags;
5. compile each ISA file with only the flags it needs.

## DESCRIPTION

Add the header provider:

```text
LinkingTo: RsimdDispatch
```

This makes bundled SIMDe headers available to C code:

```c
#include <simde/x86/avx2.h>
#include <simde/arm/neon.h>
```

## Copy the scaffold

From the downstream package root:

```{r, eval = FALSE}
RsimdDispatch::use_simd_dispatch(".")
```

The copied scaffold includes:

```text
configure
configure.win
cleanup
tools/configure-simd-dispatch.sh
src/Makevars.in
src/cpu_features.c
src/simd_dispatch.c
src/kernel_scalar.c
src/kernel_avx2.c
src/kernel_avx512.c
src/kernel_neon.c
```

## Replace the demo kernel

The demo kernel has this shape:

```c
typedef size_t (*rsd_count_nonzero_fn)(const uint8_t *x, size_t n);
```

For a real package, rename the `rsd_` and `RC_` symbols and replace the kernel
signature with your own. Keep the same structure:

```text
R API wrapper        baseline flags
CPU feature checks   baseline flags
dispatch table       baseline flags
scalar kernel        baseline flags
SSE/AVX/NEON files   per-file ISA flags
```

Do not put `-mavx2`, `-mavx512*`, or `-march=native` in global package flags.
The dispatcher must remain safe on baseline CPUs.

## Build

The generated `configure` checks compiler support and writes `src/config.h` plus
`src/Makevars`. Runtime CPU support is checked later by the installed package:

```sh
R CMD INSTALL .
```
